Experiment 2a: Analysis of Root Questions, NoContext

### 2021-11-11 analysis/05_experiment/rscripts/main_analysis.Rmd
added initial files from old repo

---
title: "Experiment 1b: Analysis of Embedded Questions, Context"
author: "mcmoyer"
date: "March 30, 2021"
---

@@

```{r, include=FALSE, warning=FALSE, echo=FALSE}
d = read.csv("../data/normed.csv")
d$ModalPresent = as.factor(d$ModalPresent)
d$Wh = as.factor(d $Wh)
d$paraphrase = as.factor(d$paraphrase)
d$VerbLemma = as.factor(d$VerbLemma)
```

# Is rating modulared by Paraphrase (MS/MA), Wh and ModalPresent?
- No significant three-way interactions

@@

```{r}

# m.ms.ma.full = lmer(normed_rating ~ cParaphrase*Wh*cModalPresent + (1+cParaphrase+Wh+cModalPresent|workerid) + (1+cParaphrase|tgrep_id), data=d.ms.ma)
# saveRDS(m.ms.ma.full, "model-msma-full.rds")

@@

m.what = lmer(normed_rating ~ cModalPresent*cParaphrase + (1+cParaphrase+cModalPresent|workerid) + (1+cParaphrase|tgrep_id) + (1+cParaphrase|VerbLemma), data=d_what) 
summary(m.what)

@@

m.how = lmer(normed_rating ~ cModalPresent*cParaphrase + (1+cParaphrase+cModalPresent|workerid) + (1+cParaphrase|tgrep_id) + (1+cParaphrase|VerbLemma), data=d_how) 
summary(m.how)

@@

m.where = lmer(normed_rating ~ cModalPresent*cParaphrase + (1+cParaphrase+cModalPresent|workerid) + (1+cParaphrase|tgrep_id) + (1+cParaphrase|VerbLemma), data=d_where) 
summary(m.where)

# m.where.simple = lmer(normed_rating ~ cParaphrase*cModalPresent-cModalPresent + (1+cParaphrase+cModalPresent|workerid) + (1+cParaphrase|tgrep_id), data=d_where) 
# summary(m.where.simple)
```

@@

m.why = lmer(normed_rating ~ cModalPresent*cParaphrase + (1+cParaphrase+cModalPresent|workerid) + (1+cParaphrase|tgrep_id) + (1+cParaphrase|VerbLemma), data=d_why) 
summary(m.why)

# m.why.simple = lmer(normed_rating ~ cParaphrase*cModalPresent-cModalPresent + (1+cParaphrase+cModalPresent|workerid) + (1+cParaphrase|tgrep_id), data=d_why) 
# summary(m.why.simple)
```

@@

m.who = lmer(normed_rating ~ cModalPresent*cParaphrase + (1+cParaphrase|workerid) + (1+cParaphrase|tgrep_id) + (1+cParaphrase|VerbLemma), data=d_who) 
summary(m.who)

# m.who.simple = lmer(normed_rating ~ cParaphrase*cModalPresent-cModalPresent + (1+cParaphrase|workerid) + (1+cParaphrase|tgrep_id), data=d_who) 
# summary(m.who.simple)
```

@@

m.when = lmer(normed_rating ~ cModalPresent*cParaphrase + (1+cParaphrase|tgrep_id) + (1+cParaphrase|VerbLemma), data=d_when) 
summary(m.when)

# simple effects
# m.when.simple = lmer(normed_rating ~ cParaphrase*cModalPresent-cModalPresent + (1+cParaphrase|tgrep_id), data=d_when) 
# summary(m.when.simple)
```

# Other Models
## Does paraphrase predict rating?
- No

```{r}
# m.ms.ma = lmer(normed_rating ~ cParaphrase + (1+cParaphrase|workerid) + (1+cParaphrase|tgrep_id) + (1+cParaphrase|VerbLemma), data=d.ms.ma)
# saveRDS(m.ms.ma, "model-msma.rds")

m.ms.ma <- readRDS("model-msma.rds")
summary(m.ms.ma)
# confint(m.ms.ma)
# cParaphrase -0.09734602 -0.03875422
```

## Are MS ratings modulated by linguistic form factors?
- no interaction between Wh-Word and Modality
- modal questions increase rating
- but compared to 'when', all other 'wh'-words decrease ratings
```{r}
d.ms = d %>% 
  filter(paraphrase=="a") %>% 
  mutate(cModalPresent = as.numeric(ModalPresent)-mean(as.numeric(ModalPresent))) %>%
  droplevels()
# Set the reference levels
contrasts(d.ms$Wh) = cbind("how.vs.when"=c(0,1,0,0,0,0),"what.vs.when"=c(1,0,0,0,0,0),
                "where.vs.when"=c(0,0,0,1,0,0),"who.vs.when"=c(0,0,0,0,1,0),
                "why.vs.when"=c(0,0,0,0,0,1))

# m.ms = lmer(normed_rating ~ cModalPresent*Wh + (1+Wh+cModalPresent|workerid) + (1|VerbLemma) + (1|tgrep_id), data=d.ms)
# saveRDS(m.ms, "model-ms.rds")

m.ms <- readRDS("model-ms.rds")
summary(m.ms)
# confint(m.ms)
```

## Are know-wh questions MA-biased?
```{r}
d.ms.ma.know = d.ms.ma %>% 
  filter(VerbLemma == "know") %>% 
  mutate(cParaphrase = as.numeric(paraphrase.dc)-mean(as.numeric(paraphrase.dc))) %>%
  mutate(cModalPresent = as.numeric(ModalPresent)-mean(as.numeric(ModalPresent))) %>%
  droplevels()
# Set the reference levels
contrasts(d.ms.ma.know$Wh) = cbind("how.vs.when"=c(0,1,0,0,0,0),"what.vs.when"=c(1,0,0,0,0,0),
                "where.vs.when"=c(0,0,0,1,0,0),"who.vs.when"=c(0,0,0,0,1,0),
                "why.vs.when"=c(0,0,0,0,0,1))
```

```{r}
d.ms.ma.know = d.ms.ma %>%
  filter(VerbLemma == "know") %>%
  mutate(cModalPresent = as.numeric(ModalPresent)-mean(as.numeric(ModalPresent))) %>%
  mutate(cParaphrase = as.numeric(paraphrase.dc)-mean(as.numeric(paraphrase.dc))) %>% 
  droplevels()

contrasts(d.ms.ma.know$Wh) = cbind("how.vs.when"=c(0,1,0,0,0,0),"what.vs.when"=c(1,0,0,0,0,0),
                "where.vs.when"=c(0,0,0,1,0,0),"who.vs.when"=c(0,0,0,0,1,0),
                "why.vs.when"=c(0,0,0,0,0,1))

# m.know = lmer(normed_rating ~ cModalPresent*cParaphrase*Wh + (1+cParaphrase+cModalPresent+Wh|workerid) + (1+cParaphrase|tgrep_id), data=d.ms.ma.know)
# saveRDS(m.know, "m-know.rds")

m.know <- readRDS("m-know.rds")
summary(m.know)
```
## Know-What
```{r know-what}
d_what_know = d.ms.ma.know %>% 
  filter(Wh=="what") %>% 
  mutate(cModalPresent = as.numeric(ModalPresent)-mean(as.numeric(ModalPresent))) %>% 
  mutate(cParaphrase = as.numeric(paraphrase.dc)-mean(as.numeric(paraphrase.dc))) %>% 
  droplevels()

m.what.know = lmer(normed_rating ~ cModalPresent*cParaphrase + (1+cParaphrase+cModalPresent|workerid) + (1+cParaphrase|tgrep_id), data=d_what_know) 
summary(m.what.know)

# m.what.simple = lmer(normed_rating ~ paraphrase*cModalPresent-cModalPresent + (1+paraphrase+cModalPresent|workerid) + (1+paraphrase|tgrep_id), data=d_what_know)
# summary(m.what.simple)
```

## Know-How
```{r know-how}
d_how_know = d.ms.ma.know %>% 
  filter(Wh=="how") %>% 
  mutate(cModalPresent = as.numeric(ModalPresent)-mean(as.numeric(ModalPresent))) %>% 
  mutate(cParaphrase = as.numeric(paraphrase.dc)-mean(as.numeric(paraphrase.dc))) %>% 
  droplevels()

m.how.know = lmer(normed_rating ~ cModalPresent*cParaphrase + (1+cParaphrase+cModalPresent|workerid) + (1+cParaphrase|tgrep_id), data=d_how_know) 
summary(m.how.know)

## Know-Where
```{r know-where model}
d_where_know = d.ms.ma.know %>% 
  filter(Wh=="where") %>% 
  mutate(cModalPresent = as.numeric(ModalPresent)-mean(as.numeric(ModalPresent))) %>% 
  mutate(cParaphrase = as.numeric(paraphrase.dc)-mean(as.numeric(paraphrase.dc))) %>% 
  droplevels()

m.where.know = lmer(normed_rating ~ cModalPresent*cParaphrase + (1+cParaphrase|workerid) + (1+cParaphrase|tgrep_id), data=d_where_know) 
summary(m.where.know)

# m.where.simple = lmer(normed_rating ~ cParaphrase*cModalPresent-cModalPresent + (1+cParaphrase+cModalPresent|workerid) + (1+cParaphrase|tgrep_id), data=d_where) 
# summary(m.where.simple)
```

## Know-Why
```{r know-why model}
d_why_know = d.ms.ma.know %>% 
  filter(Wh=="why") %>% 
  mutate(cModalPresent = as.numeric(ModalPresent)-mean(as.numeric(ModalPresent))) %>% 
  mutate(cParaphrase = as.numeric(paraphrase.dc)-mean(as.numeric(paraphrase.dc))) %>% 
  droplevels()

m.why.know = lmer(normed_rating ~ cModalPresent*cParaphrase + (1+cParaphrase|workerid) + (1+cParaphrase|tgrep_id), data=d_why_know) 
summary(m.why.know)

# m.why.simple = lmer(normed_rating ~ cParaphrase*cModalPresent-cModalPresent + (1+cParaphrase+cModalPresent|workerid) + (1+cParaphrase|tgrep_id), data=d_why) 
# summary(m.why.simple)
```

## Know-Who
```{r know-who model}
d_who_know = d.ms.ma.know %>% 
  filter(Wh=="who") %>% 
  mutate(cModalPresent = as.numeric(ModalPresent)-mean(as.numeric(ModalPresent))) %>% 
  mutate(cParaphrase = as.numeric(paraphrase.dc)-mean(as.numeric(paraphrase.dc))) %>% 
  droplevels()

m.who.know = lmer(normed_rating ~ cModalPresent*cParaphrase + (1+cParaphrase|workerid) + (1+cParaphrase|tgrep_id), data=d_who_know) 
summary(m.who.know)

# m.who.simple = lmer(normed_rating ~ cParaphrase*cModalPresent-cModalPresent + (1+cParaphrase|workerid) + (1+cParaphrase|tgrep_id), data=d_who) 
# summary(m.who.simple)
```

## Know-When
```{r know-when model}
# have to use less random effects structure
d_when_know = d.ms.ma.know %>% 
  filter(Wh=="when") %>% 
  mutate(cModalPresent = as.numeric(ModalPresent)-mean(as.numeric(ModalPresent))) %>% 
  mutate(cParaphrase = as.numeric(paraphrase.dc)-mean(as.numeric(paraphrase.dc))) %>% 
  droplevels()

m.when.know = lmer(normed_rating ~ cModalPresent*cParaphrase + (1+cParaphrase|tgrep_id), data=d_when_know) 
summary(m.when.know)

# simple effects
# m.when.simple = lmer(normed_rating ~ cParaphrase*cModalPresent-cModalPresent + (1+cParaphrase|tgrep_id), data=d_when) 
# summary(m.when.simple)
```

# Are MA ratings modulated by linguistic form factors?
- some differences based on WH-word, both 'who' and 'how' increase MA ratings compared to 'when'
```{r}
d.ma = d %>% 
  filter(paraphrase=="every") %>% 
  mutate(cModalPresent = as.numeric(ModalPresent)-mean(as.numeric(ModalPresent))) %>%
  droplevels()
# Set the reference levels
contrasts(d.ma$Wh) = cbind("how.vs.when"=c(0,1,0,0,0,0),"what.vs.when"=c(1,0,0,0,0,0),
                "where.vs.when"=c(0,0,0,1,0,0),"who.vs.when"=c(0,0,0,0,1,0),
                "why.vs.when"=c(0,0,0,0,0,1))

# m.ma = lmer(normed_rating ~ cModalPresent*Wh + (1+Wh+cModalPresent|workerid) + (1|VerbLemma) + (1|tgrep_id), data=d.ma)
# saveRDS(m.ma, "model-ma.rds")

m.ma <- readRDS("model-ma.rds")
summary(m.ma)
# confint(m.ma)
```

# Full Model with all three paraphrases

```{r, eval=FALSE, warning=FALSE, echo=FALSE}
# No three-way interactions
contrasts(d$Wh) = cbind("how.vs.when"=c(0,1,0,0,0,0),"what.vs.when"=c(1,0,0,0,0,0),
                "where.vs.when"=c(0,0,0,1,0,0),"who.vs.when"=c(0,0,0,0,1,0),
                "why.vs.when"=c(0,0,0,0,0,1))

@@

```{r, eval=FALSE, warning=FALSE, echo=FALSE}
# m.full = lmer(normed_rating ~ cModalPresent*Wh*paraphrase + (1+paraphrase+Wh+cModalPresent|workerid) + (1+paraphrase|VerbLemma) + (1+paraphrase|tgrep_id), data=centered,REML=FALSE) 
# saveRDS(m.full, "EQ-model-full-mvRE.rds")
m.full <- readRDS("EQ-model-full-mvRE.rds")
summary(m.full)
```

# Models broken up by Wh

>>>>

---
title: "Experiment 2a: Analysis of Root Questions, NoContext"
author: "mcmoyer"
date: "March 30, 2021"
output: html_document
---

@@

```{r, include=FALSE, warning=FALSE, echo=FALSE}
d = read.csv("../data/normed_nocontext_root.csv")

d$ModalPresent = as.factor(d$ModalPresent)
d$Wh = as.factor(d $Wh)
d$paraphrase = as.factor(d$paraphrase)
```

@@

# Is rating modulared by Paraphrase, Wh and ModalPresent?

- only 2-way: Wh and Modal Present, and Wh and Paraphrase
- No paraphrase and ModalPresent
- Bias for MS,but ModalPresent actually decreases the difference between MS and MA (Modal Questions overall higher than NoModal)

```{r}
# m.ms.ma.full = lmer(normed_rating ~ cParaphrase*Wh*cModalPresent + (1+cParaphrase+Wh+cModalPresent|workerid) + (1+cParaphrase|tgrep_id), data=d.ms.ma)
# saveRDS(m.ms.ma.full, "model-msma-full.rds")

@@

m.what = lmer(normed_rating ~ cModalPresent*cParaphrase + (1+cParaphrase+cModalPresent|workerid) + (1+cParaphrase|tgrep_id), data=d_what) 
summary(m.what)

@@

m.how = lmer(normed_rating ~ cModalPresent*cParaphrase + (1+cParaphrase+cModalPresent|workerid) + (1+cParaphrase|tgrep_id), data=d_how) 
summary(m.how)

@@

m.where = lmer(normed_rating ~ cModalPresent*cParaphrase + (1+cParaphrase+cModalPresent|workerid) + (1+cParaphrase|tgrep_id), data=d_where) 
summary(m.where)

m.where.simple = lmer(normed_rating ~ cParaphrase*cModalPresent-cModalPresent + (1+cParaphrase+cModalPresent|workerid) + (1+cParaphrase|tgrep_id), data=d_where)
summary(m.where.simple)
```

@@

m.why = lmer(normed_rating ~ cModalPresent*cParaphrase + (1+cParaphrase+cModalPresent|workerid) + (1+cParaphrase|tgrep_id), data=d_why) 
summary(m.why)

m.why.simple = lmer(normed_rating ~ cParaphrase*cModalPresent-cModalPresent + (1+cParaphrase+cModalPresent|workerid) + (1+cParaphrase|tgrep_id), data=d_why)
summary(m.why.simple)
```

@@

m.who = lmer(normed_rating ~ cModalPresent*cParaphrase + (1+cParaphrase|workerid) + (1+cParaphrase|tgrep_id), data=d_who) 
summary(m.who)

m.who.simple = lmer(normed_rating ~ cParaphrase*cModalPresent-cModalPresent + (1+cParaphrase|workerid) + (1+cParaphrase|tgrep_id), data=d_who)
summary(m.who.simple)
```

@@

m.when = lmer(normed_rating ~ cModalPresent*cParaphrase + (1+cParaphrase|tgrep_id), data=d_when) 
summary(m.when)

# simple effects
m.when.simple = lmer(normed_rating ~ cParaphrase*cModalPresent-cModalPresent + (1+cParaphrase|tgrep_id), data=d_when)
summary(m.when.simple)
```
# Is there an overall bias for MA? Does paraphrase predict the rating?
- No: going from MS to MA decreases the ratings, meaning that there is an MS b
```{r}
# m.ms.ma = lmer(normed_rating ~ cParaphrase + (1+cParaphrase|workerid) + (1+cParaphrase|tgrep_id), data=d.ms.ma)
# saveRDS(m.ms.ma, "model-msma.rds")

m.ms.ma <- readRDS("model-msma.rds")
summary(m.ms.ma)
```

# Are MS ratings modulated by linguistic form factors?
- yes, by the presence of a modal, which increases ratings
- one 2-way interaction wtih where vs when
```{r}
d.ms = d %>% 
  filter(paraphrase=="a") %>% 
  mutate(cModalPresent = as.numeric(ModalPresent)-mean(as.numeric(ModalPresent))) %>%
  droplevels()
# Set the reference levels
contrasts(d.ms$Wh) = cbind("how.vs.when"=c(0,1,0,0,0,0),"what.vs.when"=c(1,0,0,0,0,0),
                "where.vs.when"=c(0,0,0,1,0,0),"who.vs.when"=c(0,0,0,0,1,0),
                "why.vs.when"=c(0,0,0,0,0,1))

# m.ms = lmer(normed_rating ~ cModalPresent*Wh + (1+Wh+cModalPresent|workerid) + (1|tgrep_id), data=d.ms)
# saveRDS(m.ms, "model-ms.rds")

m.ms <- readRDS("model-ms.rds")
summary(m.ms)

# Are MA ratings modulated by linguistic form factors?
- signficant three-way interactions
- modals increase MA ratings too
- 
```{r}
d.ma = d %>% 
  filter(paraphrase=="every") %>% 
  mutate(cModalPresent = as.numeric(ModalPresent)-mean(as.numeric(ModalPresent))) %>%
  droplevels()
# Set the reference levels
contrasts(d.ma$Wh) = cbind("how.vs.when"=c(0,1,0,0,0,0),"what.vs.when"=c(1,0,0,0,0,0),
                "where.vs.when"=c(0,0,0,1,0,0),"who.vs.when"=c(0,0,0,0,1,0),
                "why.vs.when"=c(0,0,0,0,0,1))

# m.ma = lmer(normed_rating ~ cModalPresent*Wh + (1+Wh+cModalPresent|workerid) + (1|tgrep_id), data=d.ma)
# saveRDS(m.ma, "model-ma.rds")

m.ma <- readRDS("model-ma.rds")
summary(m.ma)
```

# Other Models

```{r, eval=FALSE, warning=FALSE, echo=FALSE}
# Set the reference levels
contrasts(d$Wh) = cbind("how.vs.when"=c(0,1,0,0,0,0),"what.vs.when"=c(1,0,0,0,0,0),
                "where.vs.when"=c(0,0,0,1,0,0),"who.vs.when"=c(0,0,0,0,1,0),
                "why.vs.when"=c(0,0,0,0,0,1))

@@

# Full Model with all paraphrases

- Significant three-way interactions with Task for 'the' compared to 'every'
- Significant two-way for 'a' versus 'every' 
```{r, eval=FALSE, warning=FALSE, echo=FALSE}
# m.full = lmer(normed_rating ~ cModalPresent*Wh*paraphrase + (1+paraphrase+Wh+cModalPresent|workerid) + (1+paraphrase|tgrep_id), data=centered,REML=FALSE)
# saveRDS(m.full, "model-nocontext-root-full.rds")

nocontext_model <- readRDS("model-nocontext-root-full.rds")
summary(nocontext_model)
```

# Models broken up by Wh for all paraphrases
